---
title: "leasing cleaning"
format: html
editor: visual
---

In this data notebook I will be analyzing GSA data that contains information regarding lease start and expiration dates, locations and representative info.

## Major Findings

-   There is around 50 leases that will expire this year between Maryland and DC, 10 have already expired.

```{r}
#load libraries
library(tidyverse)
library(janitor)
library(dplyr)
library(DT)
library(lubridate)
library(tidyr)
library(leaflet)
```

```{r}
#Cleaning
dc <- read.csv("data/dc.csv") %>% 
  clean_names()
maryland <- read.csv("data/maryland.csv") %>% 
  clean_names()
#Only leases
dc_lease <- dc %>% 
  filter(occupancy_right_desc=="Leased")
maryland_lease <- maryland %>% 
  filter(occupancy_right_desc=="Leased")
```

I am now focusing on getting rid of the repeated rows of information that is found in both data sets.

```{r}
#getting rid of unecessary rows
dc_lease_cleaned <- dc_lease %>% 
 select(-c(walk_score,transit_score,representative_url,senator1_url,senator2_url)) 
maryland_lease_cleaned <- maryland_lease %>% 
   select(-c(walk_score,transit_score,representative_url)) 
#getting rid of the blanks in the final column
dc_lease_cleaned<- dc_lease_cleaned %>%
  filter(lease_num != "")
maryland_lease_cleaned <-maryland_lease_cleaned %>% 
  filter(lease_num != "")
```

These are the cleaned Maryland and DC data tables that can be scrolled through and has a search button.

```{r}
datatable(maryland_lease_cleaned,  options = list(
  scrollX = TRUE,  # Enables horizontal scrolling
  autoWidth = TRUE # Adjusts column widths automatically
))
```

```{r}
datatable(dc_lease_cleaned,  options = list(
  scrollX = TRUE,  # Enables horizontal scrolling
  autoWidth = TRUE # Adjusts column widths automatically
))
```

```{r}
#separating the years so that I can determine when a lease is ending
maryland_expiration <- maryland_lease_cleaned %>% 
  separate(lease_expiration_date, into = c("month", "day", "year"), sep = "/")
dc_expiration <- dc_lease_cleaned %>% 
 separate(lease_expiration_date, into = c("month", "day", "year"), sep = "/")

```

```{r}
maryland_expiration <- maryland_expiration %>% 
  filter(year=="2025")
dc_expiration <- dc_expiration %>% 
  filter(year=="2025")
#There are 21 leases that will expire this year in DC, five have already expired.
#There are 28 leases that will expire in Maryland, five have already expired and one will expire by the end of this month. 
```

```{r}
#making a combined data set for map
all_leases <- bind_rows(
  dc_lease_cleaned %>% mutate(region = "DC"),
  maryland_lease_cleaned %>% mutate(region = "Maryland")
)
leaflet(all_leases) %>%
  addTiles() %>%
  addCircleMarkers(
    ~longitude, ~latitude, 
    color = ~ifelse(region == "DC", "blue", "red"), # Different colors for DC & MD
    radius = 3,
    fillOpacity = 0.7,
    popup = ~paste("Region:", region,)
  )
```

This map shows the dispersion of the leases that will expire in 2025.

```{r}
all_leases_expiration <- all_leases %>% 
  mutate(
    lease_expiration_date = mdy(lease_expiration_date),  # Ensure it's in Date format
    month = month(lease_expiration_date),
    day = day(lease_expiration_date),
    year = year(lease_expiration_date)
  )
all_leases_expiration<- all_leases_expiration %>% 
  filter(year=="2025")

leaflet(all_leases_expiration) %>%
  addTiles() %>%
  addCircleMarkers(
    ~longitude, ~latitude, 
    color = ~ifelse(region == "DC", "blue", "red"), # Different colors for DC & MD
    radius = 3,
    fillOpacity = 0.7,
    popup = ~paste("Lease expiration date:", lease_expiration_date)
  )

```

This map shows the dispersion of vacant properties.

```{r}
all_leases_vacant <- all_leases %>% 
  filter(bld_vacant_rsf != 0)
#there are 69 properies between Maryland and DC that have vacant space. Only 20 are from DC. 
leaflet(all_leases_vacant) %>%
  addTiles() %>%
  addCircleMarkers(
    ~longitude, ~latitude, 
    color = ~ifelse(region == "DC", "blue", "red"), # Different colors for DC & MD
    radius = 3,
    fillOpacity = 0.7,
    popup = ~paste("Building vacancies:", bld_vacant_rsf)
  )

```
